---
title: "Reproducible-Research-Course-project-1"
author: "D.Dashinov"
date: "10/21/2020"
output:
  pdf_document: default
  html_document: default
---

# Getting the data

First I'm going to write some code for downloading the data

```{r get data}
fileURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileURL, destfile = "data.zip")

unzip("data.zip", files = NULL, list = FALSE, overwrite = TRUE,
      junkpaths = FALSE, exdir = ".", unzip = "internal",
      setTimes = FALSE)

DataActivity <- read.csv("activity.csv")
```

# Question 1

What is mean total number of steps taken per day?

```{r steps per day}
library(dplyr)
library(ggplot2)

Q1 <- DataActivity %>% group_by(date) %>% summarise(Mean = mean(steps, na.rm = T))

qplot(Q1$Mean, geom="histogram",xlab="Total Steps",ylab="Counts",main="Total Steps Historgram")

summary(Q1)


```

# Question 2

What is the average daily activity pattern?

```{r Average daily activity pattern}

library(ggplot2)

df <- DataActivity %>% group_by(interval) %>% summarise(Mean = mean(steps, na.rm = T))

ggplot(df, aes(x = interval, y = Mean, group=1)) + 
        geom_path(aes(color = "orage")) + 
        ggtitle("Average daily activity pattern") + 
        ylab("Steps")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r 5-minute interval with the maximum number of steps}
df[which.max(df$Mean), ]$interval
```

# Question 3

Imputing missing values

```{r missing values}
sum(is.na(DataActivity$steps))

imputed_steps <- df$Mean[match(DataActivity$interval, df$interval)]
```

Creating a new dataset that is equal to the original dataset but with the missing data filled in.

```{r, substituting missing values}

activity_imputed <- transform(DataActivity, steps = ifelse(is.na(DataActivity$steps), 
                                        yes = imputed_steps, no = DataActivity$steps))
total_steps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")
```

Histogram of the total number of steps taken each day with and report the mean and median total number of steps taken per day.

```{r histogram}
hist(total_steps_imputed$daily_steps, col = "darkblue", xlab = "Total steps per day", ylim = c(0,30), main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))
```

Here is the mean

```{r}
mean(total_steps_imputed$daily_steps)
```

Here is the median

```{r}
median(total_steps_imputed$daily_steps)
```

Creating a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}

DataActivity$date <- as.Date(strptime(DataActivity$date, format="%Y-%m-%d"))

DataActivity$datetype <- sapply(DataActivity$date, function(x) {
        if (weekdays(x) == "Saturday" | weekdays(x) =="Sunday") 
        {y <- "Weekend"} else 
        {y <- "Weekday"}
        y
})

```

Plotting by weekdays and weekends

```{r last plot}
activity_by_date <- aggregate(steps~interval + datetype, DataActivity, mean, na.rm = TRUE)

activity_by_date$datetype <- as.factor(activity_by_date$datetype)
activity_by_date$interval <- as.numeric(activity_by_date$interval)

plot<- ggplot(activity_by_date, aes(x = interval, y = steps, color = datetype)) +
        geom_line() +
        labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
        facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)
```
